---
title: "Reproducible Research Assignment 1"
author: "Chuck Chekuri"
date: "3/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction

This report is an exercise in literate statistical programming using knittr. The activity monitoring file is downloaded and from the github repository and the r code does the analysis as specified in the assignment.


## Loading and preprocessing the data
```{r downloadandread, echo=TRUE}
# setwd("~/GitHub/RepData_PeerAssessment1")
data <- read.csv("activity.csv")

data$date <- as.POSIXct(strptime(data$date,'%Y-%m-%d'))


### What is mean total number of steps taken per day?

library(dplyr);library(ggplot2)
daily_steps <- data %>% filter(!is.na(steps)) %>% 
	group_by(date) %>% 
	summarize(total = sum(steps))
mean_steps <- round(mean(daily_steps$total),0)
median_steps <- round(median(daily_steps$total),0)

# 1. Make a histogram of the total number of steps taken each day

hist(daily_steps$total, breaks=15, col=4, xlab="Num of Steps", ylab="Frequency", main="Total Number of Stpes Taken Each Day")


#2. Calculate and report the **mean** and **median** total number of steps taken per day
```
- The mean total number of steps taken per day is :`r as.integer(mean_steps)`
- The median total number of steps taken per day is :`r as.integer(median_steps)`


### What is the average daily activity pattern?
``` {r }
interval_steps <- data %>%  
	filter(!is.na(steps)) %>%  
	group_by(interval) %>%  
	summarize(ave_steps = mean(steps))  
max_steps_interval <- max(interval_steps$ave_steps)  

int_w_max <- interval_steps %>%  
	     filter(ave_steps == max_steps_interval) %>%  
	    select(interval)

# 1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

plot(x=interval_steps$interval, y=as.integer(interval_steps$ave_steps), 
     type="l", col=4, xlab="5 min intervals in a day",
        ylab = "Average number of steps",
        main = "Average number of steps across all days")

int_w_max <- interval_steps %>% 
	     filter(ave_steps == max_steps_interval)

#### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```
The interval that contains the maximum number of steps is: `r int_w_max$interval` 

### Imputing missing values  
```{r}  
#### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)  

num_rows_with_NA <- as.integer(data %>% filter(is.na(steps)) %>% count)

```
Number of rows with missing values = `r num_rows_with_NA`
  
```{r}
#### 2. Devise a strategy for filling in all of the missing values in the dataset. 

# Fill all non missing rows with the average taken for all days for that interval.
na_rows_filled <- data %>% filter(is.na(steps)) %>% 
     merge(interval_steps, interval.x = interval.y) %>% 
     mutate( steps = round(ave_steps,1)) %>%
     select (steps, date, interval) 


#### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.  

non_na_rows <- data %>% filter(!is.na(steps)) 
all_rows_filled <- bind_rows(non_na_rows, na_rows_filled) %>%
	arrange(date,interval)
summary(all_rows_filled)


#### 4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

daily_steps_filled <- all_rows_filled %>% filter(!is.na(steps)) %>% 
	group_by(date) %>% 
	summarize(total = sum(steps))
mean_steps_filled <- round(mean(daily_steps_filled$total),0)
median_steps_filled <- median(daily_steps_filled$total)
hist(daily_steps_filled$total,breaks=15,col=4, xlab="Number of Steps",
     ylab="Frequency", main="Total Number of Stpes Taken Each Day")

summary(daily_steps);
summary(daily_steps_filled)
```


The difference in mean steps when missing values are filled is: `r mean_steps_filled - mean_steps`.
The difference in median steps when missing values are filled is: `r median_steps_filled - median_steps`

The mean is the same becuase the missing values were filled with the mean. The median sees a small change.

### Are there differences in activity patterns between weekdays and weekends?

```{r }
####1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
all_rows_filled$day <- as.factor(ifelse(weekdays(all_rows_filled$date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
#1. Make a panel plot containing a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was created using **simulated data**:
intvl_steps <- all_rows_filled %>%  
	filter(!is.na(steps)) %>%  
	group_by(day, interval) %>%  
	summarize(ave_steps = mean(steps))  
qplot(x=interval, y=ave_steps, data=intvl_steps, facets=day~., geom=c("line"),xlab="Interval",ylab="Number of Steps")

```

-----
